---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
This is a document describing the process of creating a basemap for the Chehalis in R. The hope is this can be used as a starting point and/or context for displays of any spatial data.

To do mapping in R requires a few packages. I have tried to reduce the number of packages used, but these ones seemed essential. Perhaps this list could be parred down more. The

```{r}
# For data analysis
library(tidyverse)

# For plotting
library(ggplot2)

# To read and manipulate spatial data
library(sf)

# To add scalebar and north arrow
# Have to install the package version from github on first install (as of 12/12/2018)
#library(devtools)
#install_github("oswaldosantos/ggsn") 
library(ggsn)
```


The next step is to load the various layers needed. All of the layers are shapefiles located on the computer. I tried playing around with the *maps* package to load in a basemap, but found not were the maps low resolution but they came in a format that was hard to deal with. I use the bounding box *st_bbox()* function to clip the state of washington to the Chehalis. 


```{r}
# NOAA subbasins
sub <- st_read(dsn = 'D:/Caleb_data/Chehalis/Spatial_model_2/spatial-model/Inputs.gdb', layer = 'NOAA_subbasins_w_fp') # Changed the location of this file since it no longer exists in hab folder


# Bounding box for the basemap data
buffer <- 1e4 # buffer of X meters
bbox <- st_bbox(sub) + c(-buffer, -buffer, buffer, buffer) # subtract from the xmin and ymin, add to the xmax and ymax

# Path to WA state shapefile
wa_state <- '//nwcfile/FE/watershed/Chehalis/Data/Map files/washington.shp'

# State of Washington -- land only
wa <- st_read(wa_state) %>%
  filter(CARTO_SYMB == 1) %>% # This field means 'land'
  st_transform(crs = st_crs(sub)) %>% # Reproject into the coordinate refrence system of the layer 'sub' (NAD83 Zone 10)
  st_crop(bbox)

# State of Washington -- water only
water <- st_read(wa_state) %>%
  filter(CARTO_SYMB == 4) %>%
  st_transform(st_crs(sub)) %>%  
  st_crop(bbox)

```


Now put it all together to create a basemap. 

```{r}
basemap <- ggplot() +
  theme_void() +
  geom_sf(data = water, fill = 'steelblue3') +
  geom_sf(data = wa, fill = 'white') +
  geom_sf(data = sub, fill = 'grey80', color = 'grey50') 
  # coord_sf(datum = NA) +
  # ggsn::north(sub, symbol = 16, scale = 0.15, location = 'topleft') +
  # ggsn::scalebar(data = sub, dist = 25, location = "bottomright", dd2km = F, st.size = 4)

basemap
```


Now add the streamline, displaying Fall Chinook spawning distribution

```{r}

fl <- st_read(dsn = 'D:/Caleb_data/Chehalis/Spatial_model_2/spatial-model/Outputs.gdb', layer = "flowline") %>%
  gather(species, presence, cohospawn:steelspawn) %>%
  filter(presence == 'Yes')

basemap +
  geom_sf(data = fl %>%
            filter(species == 'EDT_FChino'),
          color = 'blue') +
  coord_sf(datum = NA) # Remove gridlines. Must come after last geom_sf()

```
 Try it with a facet wrap. Takes ~10 min to render...
 
```{r} 
basemap +
  geom_sf(data = fl, color = 'blue') +
  facet_wrap(~species) +
  coord_sf(datum = NA) # Remove gridlines. Must come after last geom_sf()

#ggsave("spawner_distribution_all_species_20190205.jpg", dpi = 300, width = 15, height = 12)
```
 
Adding in new plots for temperature data
```{r}
fl_temps <- fl %>%
  select(-curr_temp, -hist_temp) %>%
  left_join(., flowline %>% 
              select(noaaid, temp_diff_2040, temp_diff_2080, curr_temp, hist_temp), by = "noaaid") %>%
  filter(presence == "Yes") %>%
  gather(year, temp_chg, temp_diff_2040:temp_diff_2080) %>%
  mutate(cc_impact = ifelse(year == "temp_diff_2040",
                            1.8,
                            3),
         year = ifelse(year == "temp_diff_2040",
                       "chg_2040",
                       "chg_2080"))

temps_no_action <- fl_temps %>%
  mutate(temp_chg = ifelse(can_ang > 170,
                           cc_impact,
                           temp_chg)) %>%
  mutate(temperature_change = cut(temp_chg,c(-Inf, -2.5, -1.5, -.5, .5, 1.5, 2.5, Inf), na.rm = T)) %>%
  select(temp_chg, temperature_change, year)

temps_scenario_1 <- fl_temps %>%
  mutate(temp_chg = ifelse(GSU %in% shade_gsu_scenario_1,
                           temp_chg,
                           ifelse(can_ang > 170,
                                  cc_impact,
                                  temp_chg)),
         temperature_change = cut(temp_chg,c(-Inf, -2.5, -1.5, -.5, .5, 1.5, 2.5, Inf), na.rm = T)) %>%
  select(temp_chg, temperature_change, year)
  
temps_scenario_2 <- fl_temps %>%
  mutate(temp_chg = ifelse(GSU %in% shade_gsu_scenario_2,
                           temp_chg,
                           ifelse(can_ang > 170,
                                  cc_impact,
                                  temp_chg)),
         temperature_change = cut(temp_chg,c(-Inf, -2.5, -1.5, -.5, .5, 1.5, 2.5, Inf), na.rm = T)) %>%
  select(temp_chg, temperature_change, year) 
  
  temps_scenario_3 <- fl_temps %>%
  mutate(temp_chg = ifelse(GSU %in% shade_gsu_scenario_3,
                           temp_chg,
                           ifelse(can_ang > 170,
                                  cc_impact,
                                  temp_chg)),
         temperature_change = cut(temp_chg,c(-Inf, -2.5, -1.5, -.5, .5, 1.5, 2.5, Inf), na.rm = T)) %>%
  select(temp_chg, temperature_change, year)
  
  
  
# group_by(chg_2040) %>%
# summarize(length = sum(Shape_Le_2, na.rm = T)) %>%
# mutate(total_length = sum(length, na.rm = T),
#        perc = (length/total_length) * 100)
```
Plot of 2040 temperature change no action

```{r}
basemap +
  geom_sf(data = temps_no_action %>%
            filter(year == "chg_2040"),
          aes(color = temperature_change), lwd = 1.5) +
  scale_colour_manual(values = c("navy", "blue", "deepskyblue", "azure", "yellow", "orange", "red")) +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("temperature_change_2040_no_action.jpg", dpi = 300, width = 15, height = 12)
  
  
```
Plot of 2080 temperature change no action

```{r}
basemap +
  geom_sf(data = temps_no_action %>%
            filter(year == "chg_2080"),
          aes(color = temperature_change), lwd = 2) +
  scale_colour_manual(values = c("navy", "blue", "deepskyblue", "azure", "yellow", "orange", "red")) +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("temperature_change_2080_no_action.jpg", dpi = 300, width = 15, height = 12)

  
  
```

Plot of 2040 temperature change scenario 1
```{r}
basemap +
  geom_sf(data = temps_scenario_1 %>%
            filter(year == "chg_2040"),
          aes(color = temperature_change), lwd = 1.5) +
  scale_colour_manual(values = c("navy", "blue", "deepskyblue", "azure", "yellow", "orange", "red")) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.position = c(.8, .7),
        legend.background = element_rect(color = "black", fill = "grey", size = 1))
  coord_sf(datum = NA)

ggsave("temperature_change_2040_scenario_1.jpg", dpi = 300, width = 15, height = 12)
  
  
```
Plot of 2080 temperature change scenario 1

```{r}
basemap +
  geom_sf(data = temps_scenario_1 %>%
            filter(year == "chg_2080"),
          aes(color = temperature_change), lwd = 2) +
  scale_colour_manual(values = c("navy", "blue", "deepskyblue", "azure", "yellow", "orange", "red")) +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("temperature_change_2080_scenario_1.jpg", dpi = 300, width = 15, height = 12)
```

Plot of 2040 temperature change scenario 2
```{r}
basemap +
  geom_sf(data = temps_scenario_2 %>%
            filter(year == "chg_2040"),
          aes(color = temperature_change), lwd = 1.5) +
  scale_colour_manual(values = c("navy", "blue", "deepskyblue", "azure", "yellow", "orange", "red")) +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("temperature_change_2040_scenario_2.jpg", dpi = 300, width = 15, height = 12)
  
  
```
Plot of 2080 temperature change scenario 2

```{r}
basemap +
  geom_sf(data = temps_scenario_2 %>%
            filter(year == "chg_2080"),
          aes(color = temperature_change), lwd = 2) +
  scale_colour_manual(values = c("navy", "blue", "deepskyblue", "azure", "yellow", "orange", "red")) +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("temperature_change_2080_scenario_2.jpg", dpi = 300, width = 15, height = 12)
```

Plot of 2040 temperature change scenario 3
```{r}
basemap +
  geom_sf(data = temps_scenario_3 %>%
            filter(year == "chg_2040"),
          aes(color = temperature_change), lwd = 1.5) +
  scale_colour_manual(values = c("navy", "blue", "deepskyblue", "azure", "yellow", "orange", "red")) +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("temperature_change_2040_scenario_3.jpg", dpi = 300, width = 15, height = 12)
  
  
```
Plot of 2080 temperature change scenario 3

```{r}
basemap +
  geom_sf(data = temps_scenario_3 %>%
            filter(year == "chg_2080"),
          aes(color = temperature_change), lwd = 2) +
  scale_colour_manual(values = c("navy", "blue", "deepskyblue", "azure", "yellow", "orange", "red")) +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("temperature_change_2080_scenario_3.jpg", dpi = 300, width = 15, height = 12)
```



```{r}
fl_temp_non_future <- fl %>%
  select(-curr_temp, -hist_temp) %>%
  left_join(., flowline %>% 
              select(noaaid, curr_temp, hist_temp, temp_diff), by = "noaaid") %>%
  filter(presence == "Yes") 

curr_temps <- fl_temp_non_future %>%
    mutate(Temperature = cut(curr_temp, c(-Inf, 18, 21, 24, Inf), na.rm = T)) %>%
    select(Temperature, curr_temp)

hist_temps <- fl_temp_non_future %>%
  mutate(temp = cut(hist_temp, c(-Inf, 18, 21, 24, Inf), na.rm = T)) %>% 
  select(temp, curr_temp)

temp_chg <- fl_temp_non_future %>%
  mutate(Temperature_Change = cut(temp_diff, c(-Inf, .5, 2, 3.5, Inf), na.rm = T)) %>%
  select(Temperature_Change, temp_diff)

curr_canang <- fl_temp_non_future %>%
  mutate(ang = cut(can_ang, c(-Inf, 20, 50, 100, 150, Inf), na.rm = T)) %>%
  select(ang, can_ang)

hist_canang <- fl_temp_non_future %>%
  mutate(ang = cut(hist_ang, c(-Inf, 20, 50, 100, 150, Inf), na.rm = T)) %>%
  select(ang, hist_ang)

canang_chg <- fl_temp_non_future %>%
  mutate(diff = can_ang - hist_ang,
         opening_angle_change = cut(diff, c(-Inf, 20, 40, 60, 100, Inf), na.rm = T)) %>%
  select(opening_angle_change, diff)
```
Plot of current temperature
```{r}
basemap +
  geom_sf(data = curr_temps, aes(color = Temperature), lwd = 1.5, show.legend = "line") +
  scale_colour_manual(values = c("blue", "yellow", "orange", "red")) +
  # geom_sf_label(show.legend = NA, fun.geometry = "line") +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.position = c(.85, .85),
        # legend.box.margin = margin(30, 10, 10, 10),
        legend.background = element_rect(fill = "grey", size = .5, linetype = "solid", colour = "black"))+
  coord_sf(datum = NA)

ggsave("current_temperature.jpg", dpi = 300, width = 15, height = 12)
```
Plot of historical temperature

```{r}
basemap +
  geom_sf(data = hist_temps, aes(color = temp), lwd = 2) +
  scale_colour_manual(values = c("blue", "orange", "red")) +
  # geom_sf_label(show.legend = NA, fun.geometry = "line") + 
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("historical_temperature.jpg", dpi = 300, width = 15, height = 12)
```

Plot of temperature difference from historical to Current

```{r}
basemap +
  geom_sf(data = temp_chg, aes(color = Temperature_Change), lwd = 1.5, show.legend = "line") +
  scale_colour_manual(values = c("blue", "yellow", "orange", "red")) +
  # geom_sf_label(show.legend = NA, fun.geometry = "line") +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.position = c(.85, .85),
        # legend.box.margin = margin(30, 10, 10, 10),
        legend.background = element_rect(fill = "white", size = .5, linetype = "solid", colour = "black"))+
  coord_sf(datum = NA)

ggsave("temperature_change.jpg", dpi = 300, width = 15, height = 12)
```

Plot of canopy opening angle
```{r}
basemap +
  geom_sf(data = curr_canang, aes(color = ang), lwd = 2) +
  scale_colour_manual(values = c("navy", "deepskyblue", "yellow", "orange", "red")) +
  # geom_sf_label(show.legend = NA, fun.geometry = "line") + 
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("current_canopy_angle.jpg", dpi = 300, width = 15, height = 12)
```

Plot of hist canopy opening angle
```{r}
basemap +
  geom_sf(data = hist_canang, aes(color = ang), lwd = 2) +
  scale_colour_manual(values = c("navy", "deepskyblue", "yellow", "orange", "red")) +
  # geom_sf_label(show.legend = NA, fun.geometry = "line") + 
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20)) +
  coord_sf(datum = NA)

ggsave("historical_canopy_angle.jpg", dpi = 300, width = 15, height = 12)
```

Plot of canopy opening angle change
```{r}
basemap +
  geom_sf(data = canang_chg, aes(color = opening_angle_change), lwd = 1.5, show.legend = "line") +
  scale_colour_manual(values = c("navyblue", "blue", "yellow", "orange", "red")) +
  # geom_sf_label(show.legend = NA, fun.geometry = "line") +
  theme(legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.position = c(.85, .85),
        # legend.box.margin = margin(30, 10, 10, 10),
        legend.background = element_rect(fill = "grey", size = .5, linetype = "solid", colour = "black"))+
  coord_sf(datum = NA)

ggsave("opening_angle_change", dpi = 300, width = 15, height = 12)
```
