---
title: "Temperature Paper Figures"
author: "Caleb Fogel"
date: "4/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(scales)
# library(gridExtra)
```

This document creates the datasets and figures needed for the temperature paper for submission to PLOS One.  

### Tables 1-4
Tables 1-4 are created manually.


### Table 5
Table 5 is filled manually with the data from the following table:

```{r}
# Set up run parameters
fishtype <- 'andadromous_network'
run_stochastic_eggtofry <- 'no'
sensitivity.mode <- 'no'
branch <- system(command = "git rev-parse --abbrev-ref HEAD", intern = TRUE)

get_os <- function(){
  sysinf <- Sys.info()
  if (!is.null(sysinf)){
    os <- sysinf['sysname']
    if (os == 'Darwin')
      os <- "osx"
  } else { ## mystery machine
    os <- .Platform$OS.type
    if (grepl("^darwin", R.version$os))
      os <- "osx"
    if (grepl("linux-gnu", R.version$os))
      os <- "linux"
  }
  tolower(os)
}
os <- get_os()
getwd()

# source("../../asrp/hab/R_files/0-Run_Habitat_Model.R")

table_5 <- read.csv('asrp_reach_data.csv') %>%
  mutate(temp_bin = case_when(asrp_temp >= 24 ~ 'high',
                              asrp_temp >= 18 & asrp_temp < 24 ~ 'med',
                              TRUE ~ 'low')) %>%
  filter(Scenario_num %in% c('cc_only', 'rip_and_climate', 'fp_temp', 'rip_and_flp', 'Current', 'Historical')) %>%
  group_by(Scenario_num, temp_bin, year) %>%
  tally() %>%
  unite(col = 'scenario', Scenario_num, year, sep = '_') %>%
  spread(temp_bin, n) %>%
  mutate(perc_high = high / (high + low + med) * 100,
         perc_med = med / (high + low + med) * 100,
         perc_low = low / (high + low + med) * 100) %>%
  select(scenario, perc_low, perc_med, perc_high)

table_5

write.csv(table_5, 'figures/table_5_temp_bins.csv')
```

### Figure 1
Created manually in ArcGIS.  It is stored in "//nwcfile/FE/watershed/Chehalis/Reports/Temperature Paper/Submission Figures/fig1_study_area.  An ArcMAP package is compressed and stored with map and all components included.


### Figure 2
Created manually in ArcGIS.  It is stored in 


### Figure 3
Figure 3 is created in "C:/caleb/asrp/asrp/docs/new_thermalscape_data/new_thermalscape_conversions.Rmd"

### Figure 4
```{r}
bowerman <- function(t,
                        phos = 0,
                        b0 = -9.053,
                        b1 = .387,
                        b2 = .029) {
  1 - plogis(logit_psm <- b0 + b1 * t + b2 * phos)
}
coho <- function(t){
                 ifelse(t < 17,
           1,
           ifelse(t >= 17 & t < 28,
                  -(1/11) * t + 2.54,
                  0))
}
chinook <- function(t){
  ifelse(t < (18/.98),
         1,
         ifelse(t >= (18/.98) & t <(24/.98),
                -(1/(6/.98)) * t + (4),
                0))
}
sthd <- function(t) {
  ifelse(t <= 17,
           t/t,
           (97.8846 / (1 + exp(-((t - 24.3522) / -.5033)))) / 100 )
}

ggplot(data.frame(x = c(12, 32)), aes(x = x)) +
  stat_function(fun = coho,aes(color = 'Coho juvenile\n summer rearing', linetype = 'coho juvenile summer rearing'), size = 1)+
  stat_function(fun = chinook,  aes(color = 'Chinook salmon juvenile\n outmigration', linetype ='chinook juvenile outmigration'), size = 1)+
  stat_function(fun = sthd,  aes(color = 'Steelhead juvenile\n summer rearing', linetype = 'steelhead juvenile summer rearing'), size = 1)+
  stat_function(fun = bowerman, aes(color = 'Spring-run Chinook salmon\n prespawning', linetype = 'spring-run Chinook prespawning'), size = 1)+
  theme_bw()+
  xlab('Temperature (7-DADM, in °C)')+
  ylab('Capacity/Productivity Multiplier')+
  scale_linetype_manual(values = c('solid', 'dashed', 'dotdash', 'dotted'), 
                        guide = FALSE) +
  scale_color_manual(
                     values = c('firebrick1', 'maroon', 'darkorange4', 'navy'),
                     guide = guide_legend(override.aes = list(
                       linetype = c('solid', 'dashed', 'dotdash', 'dotted'))))+
  theme(legend.position = c(.2, .25),
        legend.key.size = unit(1.25, 'cm'),
        legend.text = element_text(size = 14)
        )+#c(.88, .92)) +
  # theme(legend.background = element_rect(fill = 'white',
                                         # size = .25, linetype = 'solid', colour = 'black'))+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 12))+
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

ggsave('figures/fig4_temp_funcs.tiff', width = 8, height = 6, dpi = 300, compression = 'lzw')
```


### Figure 5 - 7
Figures 5-7 are created in "C:/caleb/asrp/asrp/temp_paper/create_temp_figures.Rmd.Rmd"

### Figure 8
Figure 8 is created when the model is run for fishtype == "all_species".  The code that creates this figure is in "C:/caleb/asrp/asrp/temp_paper/spawner_plots.R".  This script also creates a csv with percent change in spawner abundance from Current for each scenario and era.  This csv is used to populate some of the text in the paper. 

### Figure 9
Figure 9 is created using "C:/caleb/asrp/asrp/temp_paper/fig9_edr.Rmd".  Before figure 9 is created, the model should be run with fishtype == "all_species".


