---
title: "Exploring the 'flip'"
output: 
    html_notebook:
      toc: yes
      toc_float: true
      collapsed: false
---

In April, the spring chinook results were ~600 fish in the Newaukum and ~50 in Skookumchuck. As the observed population in Skookumchuck is ~700, we decided to remove the roads above the Skookumchuck dam, which would improve fine sediment and egg to fry survival. After doing this, the egg to fry survival in Skoookumchuck went from ~0.2 to ~0.3, while all parameters in the Newaukum stayed the same. The final spawner results, however, flipped, so that Skookumchuck had ~700 fish and the Newaukum had ~25.

This notebook explores the mathematics behind the flip. 


# Setup the lifecycle
We can explore this using a very simple 2 stage lifecycle

1. Spawner to fry -- density independent
2. Fry to spawner -- density dependent
3. Smolt to adult -- density independent

As the SAR will not impact the ratio of the two populations, we can probably leave it out, but for realism I am leaving it in for now. 

```{r}
# Lifecycle
lc <- function(spawners) {
  
  # Spawner to fry
  fry <- spawners * sp_to_fry
  
  # Assume all fry move to MS
  fry_comb <- sum(fry) # total number of movers
  fry_prop <- fry/fry_comb # proportion of total from each basin
  
  # Fry to smolt -- BH in shared MS basin
  smolts <- (fry_comb  * p_fry_to_smolt) / (1 + (p_fry_to_smolt / c_smolt) * fry_comb)
  
  # fish in ms reallocated to natal basin
  smolts_reallocated <- smolts * fry_prop 
  
  # SAR
  adults <- smolts_reallocated * sar
  
  return(adults)
}
```


# Run the lifecycle

Set up some general initial conditions

```{r}
spawners_init <- c(1000,1000) # initial pop
sp_to_fry <- c(2900, 3000) # spawner to fry, in two natal basins
p_fry_to_smolt <- 0.012 # p, fry to smolt, in shared MS basin
c_smolt <- 1000 # c, fry to smolt, in shared MS basin
sar <- 0.15 # SAR
```

Run the model across 100 generations

```{r}
spawners <- spawners_init
for (i in 1:100) {
  spawners <- lc(spawners)
}

spawners
```

Percent of the total population represented by subbasin 2 after 100 years.
```{r}
spawners[2]/sum(spawners)
```

## Modify mainstem capacity

Try running it with 100x capacity in the mainstem
```{r}
c_smolt <- 1000 * 100

spawners <- spawners_init
for (i in 1:100) {
  spawners <- lc(spawners)
}

spawners[2]/sum(spawners)
```

## Modify mainstem productivity

Try running it with 10x productivity in the mainstem
```{r}
p_fry_to_smolt <- 0.012 * 10

spawners <- spawners_init
for (i in 1:100) {
  spawners <- lc(spawners)
}

spawners[2]/sum(spawners)
```

## Modify SAR

Try changing the SAR by 3x

```{r}
sar <- 0.15 * 3

spawners <- spawners_init
for (i in 1:100) {
  spawners <- lc(spawners)
}

spawners[2]/sum(spawners)
```

## Modify the initial population

Try changing the initial pop. Change it from 1000 to 10. 

```{r}
spawners_init <- c(10,10)

spawners <- spawners_init
for (i in 1:100) {
  spawners <- lc(spawners)
}

spawners[2]/sum(spawners)
```

## Modify spawner to fry in Subbasin #1

Try changing the spawner to fry survival in Subbasin #1 by 2%. This is the only parameter which is subbasin specific. 

```{r}
sp_to_fry <- c(2900, 3000) * c(1.02, 1)

spawners <- spawners_init
for (i in 1:100) {
  spawners <- lc(spawners)
}

spawners[2]/sum(spawners)
```


# Simpliest comingling example
Lets examine the simpliest possible example of comingling at a density dependent step. This will be a two stage lifecycle, with a density independent adult to juvenile stage, and a density dependent juvenile to adult stage. Two populations will have seperate adults to juvenile parameters, but will share the juvenile to adult productivity and capacity. 


## Setup the lifecycle



```{r}

lc_simple <- function(
  adults_init = c(500, 500),
  p_a2j = c(2900, 3000),
  p_j2a = 0.012,
  c_adult = 1000,
  years = 100) {
  
  n <- matrix(NA, 
            nrow = length(adults_init), 
            ncol = years,
            dimnames = list(paste0('basin',1:length(adults_init)), 1:years)
              )
  
  adults <- adults_init
  
  for (y in 1:length(adults_init)) {
    n[y, 1] <- adults[y]
  }
  
  for (i in 2:100) {
    
    # Adult to juvenile stage - DI
    juv <- adults * p_a2j
    
    # Mix juveniles together
    juv_comb <- sum(juv) # total number of movers
    juv_prop <- juv/juv_comb # proportion of total from each basin
    
    # Juvenile to adult stage - BH
    adults_comb <- (juv_comb  * p_j2a) / (1 + (p_j2a / c_adult) * juv_comb)
    
    adults <- adults_comb * juv_prop
    
    for (y in 1:length(adults_init)) {
      n[y, i] <- adults[y]
    }
  }  
  
  return(n)
}
```

## Two basins -- p = 2900, 3000



```{r}
library(tidyverse)

lc_simple() %>%
  as.data.frame.table() %>%
  mutate(Var2 = as.numeric(Var2)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(Var2, Freq, color = Var1))

```

## Two basins -- p = 2000, 3000

```{r}

lc_simple(p_a2j = c(2000, 3000)) %>%
  as.data.frame.table() %>%
  mutate(Var2 = as.numeric(Var2)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(Var2, Freq, color = Var1))

```

## Two basins -- p = 2990, 3000

```{r}

lc_simple(p_a2j = c(2990, 3000)) %>%
  as.data.frame.table() %>%
  mutate(Var2 = as.numeric(Var2)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(Var2, Freq, color = Var1))

```

## Three basins -- p = 2800, 2900, 3000

```{r}

lc_simple(adults_init = rep(500, 3),
          p_a2j = c(2800, 2900, 3000)) %>%
  as.data.frame.table() %>%
  mutate(Var2 = as.numeric(Var2)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(Var2, Freq, color = Var1))

```

## Ten basins -- p = 2100 to 3000, by 100

```{r}

lc_simple(adults_init = rep(500, 10),
          p_a2j = seq(2100,3000,100)) %>%
  as.data.frame.table() %>%
  mutate(Var2 = as.numeric(Var2)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(Var2, Freq, color = Var1))

```

## Fifty basins -- p = 2020 to 3000, by 20

```{r}

lc_simple(adults_init = rep(500, 50),
          p_a2j = seq(2020,3000,20)) %>%
  as.data.frame.table() %>%
  mutate(Var2 = as.numeric(Var2)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(Var2, Freq, color = Var1)) +
  guides(color = FALSE)

```

# Add in natal rearing fish

Perhaps we are not seeing any stablization because we don't have natal rearing fish to help compensate?

```{r}
lc_natal <- function(
  adults_init = c(500, 500),
  p_a2j = c(2900, 3000),
  p_j2a = 0.012,
  c_adult = 1000,
  years = 100,
  prop_ds = 0.5) {
  
  n <- array(
    NA,
    c(
      years, # 100 years
      3, # natal and non natal
      length(adults_init) # number of basins
    ),
    dimnames = list(1:years, c('natal','non-natal','total'), paste0('basin',1:length(adults_init)))
  ) 
  
  adults <- adults_init
  
  for (y in 1:length(adults_init)) {
    n[1,'natal',y] <- adults[y]
  }
  
  for (i in 2:100) {
    
    # Adult to juvenile stage - DI
    juv <- adults * p_a2j
    
    # Mix juveniles together
    movers <- juv * prop_ds # number of fish moving down to MS per subbasin
    movers_comb <- sum(movers) # total number of movers
    movers_prop <- movers/movers_comb # proportion of total from each basin
    
    natal <- juv - movers
    
    juv_comb <- c(natal, movers_comb)
    
    # Juvenile to adult stage - BH
    adults_comb <- (juv_comb  * p_j2a) / (1 + (p_j2a / c_adult) * juv_comb)
    
    redist <- adults_comb[length(juv_comb)] * movers_prop
    adults <- adults_comb[-length(juv_comb)] + redist
    
    for (y in 1:length(adults_comb[-length(juv_comb)])) {
      n[i, 'natal' ,y] <- adults_comb[y]
      n[i, 'non-natal' ,y] <- redist[y]
      n[i, 'total' ,y] <- adults[y]
      
    }
  }  
  
  return(n)
}
```


## Two basins -- p = 2900 and 3000, 50% natal

```{r}
lc_natal() %>%
  as.data.frame.table() %>%
  mutate(iteration = as.numeric(Var1)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(iteration, Freq, color = Var3)) +
  facet_wrap(~Var2, scales = 'free_y')
```

## Two basins -- p = 2000 and 3000, 50% natal

```{r}
lc_natal(p_a2j = c(2000,3000)) %>%
  as.data.frame.table() %>%
  mutate(iteration = as.numeric(Var1)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(iteration, Freq, color = Var3)) +
  facet_wrap(~Var2, scales = 'free_y')
```


## 3 basins -- p = 2000, 2500, 3000, 50% natal

```{r}
lc_natal(adults_init = rep(500, 3),
          p_a2j = c(2000, 2500, 3000)) %>%
  as.data.frame.table() %>%
  mutate(iteration = as.numeric(Var1)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(iteration, Freq, color = Var3)) +
  facet_wrap(~Var2, scales = 'free_y')
```

## 10 basins -- p = 2100 to 3000, 50% natal

```{r}
lc_natal(adults_init = rep(500, 10),
          p_a2j = seq(2100,3000,100)) %>%
  as.data.frame.table() %>%
  mutate(iteration = as.numeric(Var1)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(iteration, Freq, color = Var3)) +
  facet_wrap(~Var2, scales = 'free_y')
```


## 3 basins -- p = 2000, 2500, 3000, 10% natal

```{r}
lc_natal(adults_init = rep(500, 3),
         p_a2j = c(2000, 2500, 3000),
         prop_ds = .9) %>%
  as.data.frame.table() %>%
  mutate(iteration = as.numeric(Var1)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(iteration, Freq, color = Var3)) +
  facet_wrap(~Var2, scales = 'free_y')
```



## 3 basins -- p = 2000, 2500, 3000, 5% natal

```{r}
lc_natal(adults_init = rep(500, 3),
         p_a2j = c(2000, 2500, 3000),
         prop_ds = 0.01) %>%
  as.data.frame.table() %>%
  mutate(iteration = as.numeric(Var1)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(iteration, Freq, color = Var3)) +
  facet_wrap(~Var2, scales = 'free_y')
```


# Lifecycle with density dependence before comingling
On 6/10 at a small group meeting, Larry showed his simulator which appeared to stabilize with two populations parameterized differently. The above work I did showed that any difference in those two populations created a situation where one would go extinct. Perhaps Larry's example was stabilizing because of density dependence before the comingling step? 


## Setup the lifecycle


```{r}

lc_dd <- function(
  adults_init = c(500, 500),
  p_a2j = c(1000,900),
  c_j = c(1e5, 1e5),
  p_j2a = 0.012,
  c_adult = 1000,
  years = 3000) {
  
  n <- matrix(NA, 
            nrow = length(adults_init), 
            ncol = years,
            dimnames = list(paste0('basin',1:length(adults_init)), 1:years)
              )
  
  adults <- adults_init
  
  for (y in 1:length(adults_init)) {
    n[y, 1] <- adults[y]
  }
  
  for (i in 2:years) {
    
    # Adult to juvenile stage - DI
    juv <- (adults  * p_a2j) / (1 + (p_a2j / c_j) * adults)
    
    # Mix juveniles together
    juv_comb <- sum(juv) # total number of movers
    juv_prop <- juv/juv_comb # proportion of total from each basin
    
    # Juvenile to adult stage - BH
    adults_comb <- (juv_comb  * p_j2a) / (1 + (p_j2a / c_adult) * juv_comb)
    
    adults <- adults_comb * juv_prop
    
    for (y in 1:length(adults_init)) {
      n[y, i] <- adults[y]
    }
  }  
  
  return(
    n %>%
      as.data.frame.table() %>%
      mutate(Var2 = as.numeric(Var2)) %>%
      ggplot +
      theme_bw() +
      geom_line(aes(Var2, Freq, color = Var1))
         )
}

```


```{r}
lc_dd(p_a2j = c(1000,1000),
      c_j = c(3e7, 1e7)) +
  lims(y = c(0,1000))
```


# Add in natal rearing fish with density dependent movers


```{r}
lc_ddmovers <- function(
  adults_init = c(500, 500),
  p_a2j = c(650, 800),
  c_j = c(70000, 40000),
  p_j2a = 0.012,
  c_adult = c(70000, 40000, 150000),
  years = 100,
  prop_ds = 0.5) {
  
  n <- array(
    NA,
    c(
      years, # 100 years
      3, # natal and non natal
      length(adults_init) # number of basins
    ),
    dimnames = list(1:years, c('natal','non-natal','total'), paste0('basin',1:length(adults_init)))
  ) 
  
  adults <- adults_init
  
  for (y in 1:length(adults_init)) {
    n[1,'natal',y] <- adults[y]
  }
  
  for (i in 2:100) {
    
    # Adult to juvenile stage - DI
    juv <- (adults * p_a2j)/(1 + (p_a2j/c_j)*adults)
    
    # Mix juveniles together
    movers <- (adults * p_a2j) - juv # DD movers,number of fish moving down to MS per subbasin
    movers_comb <- sum(movers) # total number of movers
    movers_prop <- movers/movers_comb # proportion of total from each basin
    
    juv_comb <- c(juv, movers_comb)
    
    # Juvenile to adult stage - BH
    adults_comb <- (juv_comb  * p_j2a) / (1 + (p_j2a / c_adult) * juv_comb)
    
    redist <- adults_comb[length(juv_comb)] * movers_prop
    adults <- adults_comb[-length(juv_comb)] + redist
    
    for (y in 1:length(adults_comb[-length(juv_comb)])) {
      n[i, 'natal' ,y] <- adults_comb[y]
      n[i, 'non-natal' ,y] <- redist[y]
      n[i, 'total' ,y] <- adults[y]
      
    }
  }  
  
  return(n)
}
```

```{r}
lc_ddmovers() %>%
  as.data.frame.table() %>%
  mutate(Var1 = as.numeric(Var1)) %>%
  ggplot +
  theme_bw() +
  geom_line(aes(Var1, Freq, color = Var3)) +
  facet_wrap(~Var2)
```


