---
title: "steelhead.age.structure.and.SAR"
author: "nwfsc"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \usepackage{subfig}
- \usepackage{float}
#- \fancyhead[CO,CE]{\bf{\emph{DRAFT}}}
- \fancyhead[L]{}
- \fancyhead[R]{}
- \fancyfoot[C]{\bf{\emph{DRAFT}}}
- \fancyfoot[R]{\thepage}
- \raggedright
- \setlength{\parindent}{1em}
- \setlength{\parskip}{1em}
- \usepackage{lineno}
- \usepackage{units}
- \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 6
  html_document:
    depth: 6
    number_sections: yes
    theme: readable
    toc: yes
    toc_float:
      collapsed: yes
  word_document:
    toc: yes
---

***`r format(Sys.time(), "%Y %B %d %X %Z")`***

```{r load-tidyverse, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
```


# Examining age structure

First, we look at how closely the Chehalis winter steelhead LCM adheres to observations from fisheries data collections from the Quinault Indian Nation. Below are comparisons of ages of first-time spawning winter steelhead.

A crosswalk between definition of age terms is needed before proceeding. The QIN data names fish ages according to X.Y, where X = freshwater age and Y = saltwater age. In their data, they define that an age of 2.1 corresponds to a 2 year smolt outmigrant (2 freshwater years), followed by 1 saltwater year with migration to freshwater in winter, resulting in a 4 year old spawner. In the NOAA LCM, this would correspond to a 2.2, which is also a 4 year old spawner. Below, we translate QIN age into the NOAA LCM definition.

The winter steelhead NOAA LCM must be run first (or load a saved workspace that includes outputs from a run of the winter steelhead LCM) before this do

```{r run-steelheadLCM}
mypath <- "~/jorgensen/data/R/chehalis.LCM/chehalis.asrp/outputs/steelhead/lcm/"
load(paste0(mypath,"steelhead.age.test.11.01.2019.RData"))

```

```{r calculate-age-info, echo=FALSE}
# Age composition smolt ages in adult return

#    proportions of smolts going out of Chehalis basin, by age:
age1smolts.firstspawn <- sum(N['age1.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
age2smolts.firstspawn <- sum(N['age2.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
age3smolts.firstspawn <- sum(N['age3.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])

#    number age in adult returns:
age3.firstspawn.age1.2 <- sum(N['age3.ocean', ]) * b3 * sum(N['age1.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ]) # 100% age1 and age2 smolts
age3.firstspawn.age2.1 <- sum(N['age3.ocean', ]) * b3 * sum(N['age2.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ])# 100% age1 and age2 smolts

age4.firstspawn.age1.3 <- sum(N['age4.ocean', ]) * b4 * sum(N['age1.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
age4.firstspawn.age2.2 <- sum(N['age4.ocean', ]) * b4 * sum(N['age2.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
age4.firstspawn.age3.1 <- sum(N['age4.ocean', ]) * b4 * sum(N['age3.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
  
age5.firstspawn.age1.4 <- sum(N['age5.ocean', ]) * b5 * sum(N['age1.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
age5.firstspawn.age2.3 <- sum(N['age5.ocean', ]) * b5 * sum(N['age2.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
age5.firstspawn.age3.2 <- sum(N['age5.ocean', ]) * b5 * sum(N['age3.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])

age6.firstspawn.age1.5 <- sum(N['age6.ocean', ]) * 1 * sum(N['age1.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
age6.firstspawn.age2.4 <- sum(N['age6.ocean', ]) * 1 * sum(N['age2.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])
age6.firstspawn.age3.3 <- sum(N['age6.ocean', ]) * 1 * sum(N['age3.smolts', ]) / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ])

total.all.ages <- sum(#NA, #1.1
                      age3.firstspawn.age1.2,
                      age4.firstspawn.age1.3,
                      age5.firstspawn.age1.4,
                      age6.firstspawn.age1.5,
                      
                      age3.firstspawn.age2.1,
                      age4.firstspawn.age2.2,
                      age5.firstspawn.age2.3,
                      age6.firstspawn.age2.4,
                      
                      age4.firstspawn.age3.1,
                      age5.firstspawn.age3.2,
                      age6.firstspawn.age3.3
                      )

prop.total.all.ages <- c(  age3.firstspawn.age1.2,
                           age4.firstspawn.age1.3,
                           age5.firstspawn.age1.4,
                           age6.firstspawn.age1.5,
                           
                           age3.firstspawn.age2.1,
                           age4.firstspawn.age2.2,
                           age5.firstspawn.age2.3,
                           age6.firstspawn.age2.4,
                           
                           age4.firstspawn.age3.1,
                           age5.firstspawn.age3.2,
                           age6.firstspawn.age3.3,
                           NA# for their 3.3 (corresponding to our 3.4, we don't have that)
                         )/total.all.ages

# Checking to see if math works out right, looks like 
# total.all.ages  is the same as this below:
# sum(N[firstspawn.stages,] %>% colSums)

# Age composition from L Lestelle (QIN data):

QIN.age.comp <- c(0.007,# 1.1, 1.2
                  0.010,# 1.2, 1.3
                  0.001,# 1.3, 1.4
                  NA, # 1.4, 1.5
                  NA,# need for our 2.1
                  0.469,# 2.1, 2.2
                  0.388,# 2.2, 2.3
                  0.021,# 2.3, 2.4
                  NA,# need for our 3.1
                  0.001,# 3.1, 3.2
                  0.069,# 3.2, 3.3
                  0.032# 3.3
                  )
age.comparison <- data.frame(lcm.age = c("1.2", "1.3", "1.4", "1.5",
                                      "2.1",  
                                      "2.2", "2.3", "2.4",
                                      "3.1", 
                                      "3.2", "3.3", NA),
                             qin.age = c("1.1", "1.2", "1.3", NA,
                                          NA, 
                                          "2.1",  "2.2", "2.3",
                                          NA,
                                          "3.1", "3.2", "3.3"),
                             lcm.prop = round(prop.total.all.ages, digits=3), 
                             QIN.prop = QIN.age.comp, 
                             lcm.minus.qin = prop.total.all.ages - QIN.age.comp,
                             stringsAsFactors = FALSE)

```



```{r lineplot-age-structure, echo=FALSE, fig.cap="Line plot comparison of ages of first-time Chehalis winter steelhead observations from QIN fisheries data and from output from the NOAA LCM.", fig.align='center'}
plot(1:length(age.comparison$lcm.age),
     age.comparison$lcm.prop, type="b",
     ylab="Proportion", xlab="Age (Fresh.Salt; LCM age method)",
     ylim=c(0,0.6), axes=FALSE,
     yaxs="i", xpd=NA,
     main="Chehalis Winter steelhead\nadult return age structure (first-time spawners)")
axis(1, at=1:length(age.comparison$lcm.age), c(age.comparison$lcm.age))
axis(1, at=12, "3.4")
axis(2, las=1)
box()
points(1:length(age.comparison$lcm.age), age.comparison$QIN.prop, 
       pch=21, bg="black", xpd=NA)
text("QIN-NA", x=c(4, 5, 9), y=c(0.05, 0.02, 0.01), cex=0.6)
text("LCM-NA", x=12, y=0.02, cex=0.6)
legend("topleft",
       legend=c("QIN mean observed 1981-2019", "LCM model estimates"),
       pch=21, pt.bg=c("black", "white"), bty = "n")

```


```{r lineplot-age-structure-diff, echo=FALSE, fig.cap="Line plot differences of ages of first-time Chehalis winter steelhead observations from QIN fisheries data and from output from the NOAA LCM.", fig.align='center'}
plot(1:length(age.comparison$lcm.age), age.comparison$lcm.minus.qin,
     type="b", axes=FALSE, ylim=c(-0.1, 0.1),
     xlab="Age (Fresh.Salt; LCM age method)",
     ylab="LCM age prop. - QIN age prop.")
abline(h=0, lty=2)
axis(1, at=1:length(age.comparison$lcm.age), age.comparison$lcm.age)
axis(1, at=12, "3.4")
axis(2, las=1)
```

The following is a representation in barplot form.

```{r barplot-age-structure, echo=FALSE, fig.cap="Barplot comparison of ages of first-time Chehalis winter steelhead observations from QIN fisheries data and from output from the NOAA LCM.", fig.align='center'}
# barplot
bp1 <- barplot(t(age.comparison[,c(4,3)]), beside=TRUE, las=1,
               ylab="Proportion", xlab="Age (Fresh.Salt; LCM age method)",
               ylim=c(0,0.6), col=c("black", "white"),
               main="Chehalis Winter steelhead\nadult return age structure (first-time spawners)")
#bp1
axis(1, at=apply(bp1, 2, median), age.comparison$lcm.age)
axis(1, at=median(bp1[,12]), "3.4")
legend("topleft",
       legend=c("QIN mean observed 1981-2019", "LCM model estimates"),
       fill=c("black", "white"), bty = "n")

```


There are some ages for which each data source does not have representations. For example, QIN observations include age 3.4, corresponding to a 7 year old first-time spawner, and there is no NOAA LCM equivalent. The NOAA LCM does include 7 and 8 year olds that are re-spawners. And, the NOAA LCM includes some ages that the QIN data do not include: 1.5, 2.1, and 3.1 first-time spawners.

# SAR

```{r SAR-calcs, echo=TRUE}
# SAR
#  including kelts
(sum(N['total.run', ]) 
 / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ]))
#  excluding kelts
(sum(N[firstspawn.stages,] %>% colSums) 
  / sum(N['age1.smolts', ] + N['age2.smolts', ] + N['age3.smolts', ]))


```