---
title: "Proposal to use new prespawn function"
author: "Caleb Fogel"
date: "October 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

The `bowerman_ps` function comes from `Bowerman, 2018, Prespawn Mortality of Female Chinook Salmon Increases with Water Temperature and Percent Hatchery Origin`.  The inputs for this function are 7-DADM temperature (t, 째C) and percent hatchery origin fish (phos).  For our purposes we assume 0% hatchery fish, as we do not model hatchery origin fish in the NOAA model. This model was developed from observations within the Willamette River basin, Oregon.

First, we model prespawn mortality (psm), which is what they model in the report:

```{r}
bowerman_psm <- function(t,
                        phos = 0,
                        b0 = -9.053,
                        b1 = .387,
                        b2 = .029) {
  logit_psm <- b0 + b1 * t + b2 * phos
  psm <- plogis(logit_psm)
  plot(t, psm, type = 'l')
}
bowerman_psm(t = 12:28)
```

Next, we model prespawn survival, which is simply `1 - psm).  This is what we use to calculate prespawn survival in the NOAA model:

```{r}
bowerman_ps <- function(t,
                        phos = 0,
                        b0 = -9.053,
                        b1 = .387,
                        b2 = .029) {
  logit_pss <- b0 + b1 * t + b2 * phos
  pss <- 1 - plogis(logit_pss)
  
  plot(t, pss, type = 'l', xaxt = 'n', ylab = '')
  axis(side = 1, at = c(12:28))
  title(xlab = 'Temperature (Maximum 7-DADM, in 째C)', ylab = 'Productivity Multiplier')
  
}
bowerman_ps(t = 12:28)
```

We model both prespawn mortality and prespawn survival above from 12째 - 28째 in order to include all temperature values that we experience in the Chehalis River basin in the NOAA model

Next we plot a comparison between the Bowerman, Cramer, and Willamette functions:

```{r}
bowerman_ps.alt <- function(t,
                            phos = 0,
                            b0 = -9.053,
                            b1 = .387,
                            b2 = .029) {
  logit_pss <- b0 + b1 * t + b2 * phos
  pss <- 1- plogis(logit_pss)
  plot(t, pss, type = 'l', ylim=c(0,1), las=1)
}

bowerman_ps.alt(t = 12:28)

willamette <- function(t,
                       phos=0,
                       b0 = -13.2061,
                       b1 = 0.7919,
                       b2 = 2.0923) {
  # temps.converted <- -3.277448 + 1.011789*temps
  logit_pss <- b0 + b1 * (-3.277448 + 1.011789*t) + b2 * phos
  pss <- 1- plogis(logit_pss)
  pss
}

lines(12:28, willamette(t = 12:28), col="blue")

cramer.prespawn <- function(t){
  # cites go to Cramer 2001
  # Cramer, S. 2001. The relationship of stream habitat features to potential
  # for production of four salmonid species. S.P. Cramer & Associates, Inc.,
  # February 2001 report to Oregon Building Industry Association, Gresham, OR.
  # where x = prespawning temperature
  # Applications where it's been used in LCMs:
  # mean of daily max 15 July through 15 August; Bartz et al. 2006; Scheuerell et al. 2006
  # mean of daily max Aug. - Sept.; Jorgensen et al. 2009; Honea et al. 2009; Honea et al. 2016
  y <- rep(NA, length(t))
  for(i in 1:length(t)){
    if (t[i] < 16) {y[i] <- 1}
    if (t[i] >= 16 & t[i] < 22.6) {y[i] <- 1 - 0.15*(t[i] - 16)}
    if (t[i] >= 22.6) {y[i] <- 0.01}
  }
  y
}
lines(12:28, cramer.prespawn(t=12:28), col="purple")

```
