```{r, echo=FALSE, include=FALSE}
library(knitr)
library(tidyverse)
load('myEnvironment.RData')
```
# Chinook


# Small Stream Capacity
#### Introduction:

We begin with the SWIFD streamline layer broken into approximately 200m segments.  These segments have an assigned reach area (reach length * reach width) but area not broken into specific habitat types (pool, riffle, beaver pond).  We use the NOAA Habitat Model to separate each reach into pool, riffle, and beaver pond habitat and augment area by wood multipliers where appropriate and based on the specific habitat scenario definitions.  We do not calculate winter capacity or a temperature scalar for spring chinook

#### 1. Divide reaches into pool and riffle habitat:

Reaches are divided into pool and riffle habitat using the following pool percentages based on slope and landcover type:

```{r, echo=FALSE, results = 'asis'}
kable(ss.dist%>%spread(lc, pool.perc)%>%mutate(slope.class = ifelse(slope.class == "low", "<2%", ifelse(slope.class == "med", "2-4%", ">4%")))%>%rename(slope = slope.class)%>%arrange(match(slope, c("<2%", "2-4%", ">4%"))))
```

Summer pool and riffle areas area calculated using the following equations:

Summer habitat | Area calculation
---------------|-----------------
Pool           | reach area * percent pool
Riffle         | reach area * (1 - percent pool)


#### 2. Convert some pool and riffle area into beaver pond area:
In the Historical and Beaver Pond scenario we reallocate 15% of both the pool and riffle habitat into beaver pond habitat.  We assume 6 beaver ponds per km, and a median pond area of 500 m^2 (Pollock 2004) 

Summer habitat | Area calculation
---------------|-----------------
Pool           | reach area * percent pool * .85
Riffle         | reach area * (1 - percent pool) * .85
Beaver pond    | reach length * 3 / 10000

* *6 ponds per km and 500 m^2^ per pond translates to 3000 m^2^ per km, or 3 m^2^ per meter of river length*  
* *reach length multiplied by 3m^2^ per meter of rivre length gives us beaver pond area in m^2^*
* *Convert to hectares using 10000 m^2^ per hectare*


#### 3. Adjust area based on the addition of wood: 
We calculate a wood multiplier at the subbasin level based on bar and bank lengths within a particular subbasin:

* summer wood multiplier = ((bank length * 1.16) + bar length * 2 * 1.03)) / bank length + (bar length * 2))
    * if no bar or bank length for a subbasin, we use the mean of the wood multipliers across all subbasins, 1.11
    
We multiply area by the wood scalar in the Historical and Wood scenarios only


#### 4. Barriers

We only restrict rearing habitat above barriers if barriers have a passage rating of 0.  If the passage rating = 0, we set the area to 0.  Otherwise we do not restrict the rearing area


#### 5. Chinook scalar

When both spring and fall chinook are present in a given reach or habitat unit we use a scalar to determine how use of the habitat is distributed between spring and fall chinook

Spring Chinook present | Fall Chinook present | Spring Chinook scalar | Fall Chinook scalar
-----------------------|----------------------|-----------------------|--------------------|
Yes | Yes | .19 | .81
No | Yes | 1 | 1
Yes | No | 1 | 1
No | No | 1 | 1 

#### 6. Calculate capacity: 

Capacity = Area * Density * Chinook scalar

Density is defined as fish/ha and is decribed in the following table:

```{r, echo=FALSE, results = 'asis'}
kable(density%>%filter(Habitat %in% c("Pool", "Riffle", "Beaver.Pond"))%>%spread(Habitat, Density))
```

#### 7. Capacity review:

Scenario| Beaver ponds present? | Temperature scalar| Wood scalar? | Passage barriers present?
--------|-----------------------|-------------------|--------------|--------------------------
Current | No | None | No | Yes
Barrier | No | None | No | No
Beaver  | Yes | None | No | Yes
Fine sediment| No | None | No | Yes
Floodplain | No | None | No | Yes
Historical | Yes| None | Yes | No
LR bank | No | None | No | Yes
LR length | No | None| No | Yes
Shade | No | None | No | Yes
Wood | No | None | Yes | Yes


# Large river capacity

We begin with a layer of handmapped, attributed large river lines.  We use the NOAA Habitat Model to augment area by wood multipliers where appropriate and based on the specific habitat scenario definitions.  We do not apply a temperature scalar for spring chinook.

#### 1. Large river lengthening

We apply a lengthening scalar to large river habitat in the LR length and Historical scenarios to reflect changes in stream meander length. 

* Length scalars are applied at the reach level.  
* Where length scalars do not exist, large river area does not change.
* Under Historical and LR length scenarios, large river area = length * length scalar * width
* Large river lengthening does not affect backwater habitat

#### 2. Large river bank modification

Modified bank habitat is replaced with natural bank habitat under the LR bank and Historical scenarios.  Natural and modified bank habitats were both handmapped by Jamie, so all we do here is use one or the other depending on the scenario.

#### 3. Adjust area based on the addition of wood

We calculate a wood multiplier at the subbasin level based on bar and bank lengths within a particular subbasin:

* summer wood multiplier = ((bank length * 1.16) + bar length * 2 * 1.03)) / bank length + (bar length * 2))
    * if no bar or bank length for a subbasin, we use the mean of the wood multipliers across all subbasins, 1.11
    
We multiply area by the wood scalar in the Historical and Wood scenarios only

#### 4. Barriers

We only restrict rearing habitat above barriers if barriers have a passage rating of 0.  If the passage rating = 0, we set the area to 0.  Otherwise we do not restrict the rearing area

#### 5. Chinook scalar

When both spring and fall chinook are present in a given reach or habitat unit we use a scalar to determine how use of the habitat is distributed between spring and fall chinook

Spring Chinook present | Fall Chinook present | Spring Chinook scalar | Fall Chinook scalar
-----------------------|----------------------|-----------------------|--------------------|
Yes | Yes | .19 | .81
No | Yes | 1 | 1
Yes | No | 1 | 1
No | No | 1 | 1 

#### 5. Calculate capacity

Capacity = Area * Density * Chinook scalar

Density is defined as fish/ha and is decribed in the following table:

```{r, echo=FALSE, results = 'asis'}
kable(density%>%filter(Habitat %in% c("Bank", "HM_Bank", "Bar_boulder", "Bar_gravel", "Bar_sand"))%>%spread(Habitat, Density))
```

#### 6. Review

Scenario| Large river lengthening? | Large river bank modifications restored? | Temperature scalar | Wood scalar? | Passage barriers present?
--------|--------------------------|------------------------------------------|--------------------|--------------|--------------------------
Barrier | No | No | None | No | No
Beaver | No | No | None | No | Yes
Current | No | No | None | No | Yes
Fine Sediment | No | No | None | No | Yes
Floodplain | No | No | None | No | Yes
Historical | Yes | Yes | None | Yes | No
LR bank | No | Yes | None | No | Yes
LR length | Yes | No | None | No | Yes
Shade | No | No | None | No | Yes
Wood | No | No | None | Yes | Yes

# Floodplain capacity
#### 1. Introduction

We begin with a layer of handmapped floodplain units.  We attach attributes such as barrier presence using the attributed "flowline" streamline created in the spatial model.  We do not apply temperature or wood scalars to floodplain habitat

#### 2. Side channel habitat

* Side channel habitat is broken into side channel pools and side channel riffles based on the same slope and landcover criteria as small stream pools and riffles
* We do not convert side channel habitat to beaver pond habitat in any scenarios

#### 3. Calculate winter area

We use a winter pool scalar of .43 to convert summer pool area into winter pool area.  The remaining winter area becomes riffle area. 

Winter habitat | Area calculation
---------------|-----------------
Side channel pool           | summer side channel pool area * .43
Side channel riffle         | summer side channel riffle area + (.57 * summer side channel pool area)

#### 4. Historical side channel

We apply a side channel multiplier to select reaches of the "flowline" streamline to create historical side channel area.  
* Side channel scalars are applied at the reach level.  
* Where side channel scalars do not exist,  no new floodplain area is created.
* Under Historical and Floodplain scenarios, historical side channel length = reach length * side channel scalar * width

#### 5. Barriers

We apply only restric habitat above barriers if barriers have a passage rating of 0.  If the passage rating = 0, we set the area to 0.  Otherwise we do not restrict the rearing area.  Barriers area not applied in the Historical, Barriers, and Floodplain scenarios

#### 6. Chinook Scalar

When both spring and fall chinook are present in a given reach or habitat unit we use a scalar to determine how use of the habitat is distributed between spring and fall chinook

Spring Chinook present | Fall Chinook present | Spring Chinook scalar | Fall Chinook scalar
-----------------------|----------------------|-----------------------|--------------------|
Yes | Yes | .19 | .81
No | Yes | 1 | 1
Yes | No | 1 | 1
No | No | 1 | 1 

#### 7. Calculate Capacity

Capacity = Area * Density * chinook scalar

Density is defined as fish/ha and is decribed in the following table:

```{r, echo=FALSE, results = 'asis'}
kable(density%>%filter(Habitat %in% c("FP_Channel", "FP_Pond_lg", "FP_Pond_sm", "Lake", "Marsh", "SC_pool", "SC_riffle", "Slough_lg", "Slough_sm"))%>%spread(Habitat, Density))
```

#### 7. Review

Scenario | Hist SC addition | Temperature scalar | Wood scalar | Passage barriers present?
---------|------------------|--------------------|-------------|--------------------------
Barrier | No | None | None | No
Beaver | No | None | None | Yes
Current | No | None | None | Yes
Fine Sediment | No | None | None | Yes
Floodplain | Yes | None | None | No
Historical | Yes | None | None | Yes
LR bank | No | None | None | No
LR length | No | None | None | Yes
Shade | No | None | None | Yes
Wood | No | None | None | Yes

# Productivity
#### 1. Introduction

Subbasin productivity is calculated as a weighted average of reach level productivities in the majority of habitat scenarios.  We begin with a base survival value assigned to each habitat type.  We then calculate the value of the capacity of each reach as a percentage of the total capacity of each subbasin.  We multiply each reach's base survival by the percent capacity, and in some scenarios, a wood scalar, and then sum these values to calculate a weighted subbasin survival.

#### 2. Small stream productivity

Small stream base productivity is defined in the following table:

```{r, echo = FALSE, results = 'asis'}
kable(survival%>%filter(Habitat %in% c("Pool", "Riffle", "Beaver.Pond")) %>% gather(Wood, Base.survival, surv.base:wood.surv.base) %>% mutate(Wood = ifelse(Wood == "surv.base", "No", "Yes")) %>% spread(Habitat, Base.survival))
```

* Small stream base productivity is only defined for summer as we do not calculate a winter survival for spring chinook
* We define different small stream productivity values for scenarios with wood loading (Wood, Historical) and those without wood loading

#### 3. Large river and backwater productivity

Large river and backwater base productivity area defined in the following table:

```{r, echo = FALSE, results = 'asis'}
kable(survival%>%filter(Habitat %in% LgRiver_habs) %>% select(-wood.surv.base) %>%rename(Base.survival = surv.base) %>% spread(Habitat, Base.survival))
```

* Large river and backwater have the same base survival for current and historical wood scenarios, however large river base survival is multiplied by the subbasin-level wood multiplier in Historical and Wood scenarios

#### 4. Floodplain productivity

```{r, echo = FALSE, results = 'asis'}
kable(survival%>%filter(Habitat %in% Floodplain_habs) %>% select(-wood.surv.base) %>%rename(Base.survival = surv.base) %>% spread(Habitat, Base.survival))
```

* Floodplain productivity is not affected by wood

#### 5. Total subbasin level productivity

```{r, echo = FALSE, results = 'asis'}
surv_table <- data.frame(
Habitat = c("Bank, HM Bank, Bar, Bar-boulder, Bar-gravel, Bar-sand, Backwater", "Pool, Riffle, Beaver Pond, FP channel, lake, marsh, fp pond lg, fp pond sm, sc pool, sc riffle, slough lg, slough sm"),
Reach_survival_calculation_summer = c("Base survival * wood scalar * % of total capacity", "Base survival * % of total capacity"),
Reach_survival_calculation_winter = c(NA, NA)
)
kable(surv_table)
```

* % of total capacity = capacity of reach or habitat unit / capacity of reach

##### Total subbasin productivity is equal to the sum of all reach productivities within a subbasin

####6. Productivity special cases

We use a different method to calculate total productivity in the LR bank, LR length and Shade scenarios.  In these cases we use the following equation:

Productivity = LR survival hist ((LR bank, LR Length or shade scenario) - LR survival curr (current scenario) * % LR) + Survival of current scenario
